---
title: 'Statistical Inference Project, Part 2: Basic Inferential Data Analysis'
author: 'Tehty'
date: "Jul 23, 2015"
output: pdf_document
---
## Overview

In this second portion of the assignment, we're going to analyze the ToothGrowth data in the R datasets package.

Steps:

1.  Load the ToothGrowth data and perform some basic exploratory data analyses 
2.	Provide a basic summary of the data.
3.	Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
4.	Conclusions and the assumptions  

The `ToothGrowth` data set consists of 60 observations of 3 variables:

* `len` (numeric)
* `supp` (VC or OJ)
* `dose` (numeric)

```{r}
library(datasets)
head(ToothGrowth)
summary (ToothGrowth)

library(ggplot2)
ggplot(ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
  geom_bar(stat="identity",) +
  facet_grid(. ~ supp) +  
  labs(x ="Dose", y="Tooth Length")+ 
  guides(fill=guide_legend(title="Supplement type"))

```

As per the histogram above, it is a clear positive correlation between the tooth length and the dose levels.

```{r}
graph <- ggplot(ToothGrowth, aes(x=as.factor(dose), y=len)) 
graph + geom_point(aes(color=supp), size = 4, alpha = 1/2) + labs(title = "Tooth Growth") + labs(x ="Dose", y="Tooth Length") 

```

From above results, we can identify below

1. Tooth length with VC supplement has wider distribution than those with OJ supplement
2. Teeth are longer with OJ than those with VC at the dose 0.5 and 1.0 level
3. The larger the dose, the longer the tooth.


Now, let us examing the variance in tooth length in relation of the supplement type.

```{r}
fitting <- lm(len ~ dose + supp, ToothGrowth)
summary(fitting)
```
Assumption: All else equal

The intercept is `r fitting$coefficients[[1]]`, means that with no supplement of
Vitamin C, the average tooth length is `r fitting$coefficients[[1]]`. 

The coefficient of `dose` is `r fitting$coefficients[[2]]`. When increasing dose 1 mg, would increase the tooth length of `r fitting$coefficients[[2]]` 

The last coefficient `suppVC` with the value is `r fitting$coefficients[[3]]` explains that a given dose of VC, would result in decrease of `r abs(fitting$coefficients[[3]])` in the tooth length. 

Confidence intervals for these two variables and the intercept as below
```{r}
confint(fitting)
```

For each coefficient (the intercept, `dose` and `suppVC`), the null hypothesis is where the coefficent is 'zero' - means that no tooth length variation is explained by that variable.

All _p_-values are less than 0.05, rejecting the null hypothesis and suggesting
that each variable explains a significant portion of variability in tooth length,
with the assumption of 5% of significance level.

Hypothesis Test

Cosidering there are three levels of dose (0.5, 1.0, 2.0), we perform t test in following orders: (1) dose 0.5 vs. dose 1.0; (2) dose 1.0 vs. dose 2.0; and (3) dose 0.5 vs. dose 2.0.

```{r}

t.test(ToothGrowth$len[ToothGrowth$dose == 1.0], ToothGrowth$len[ToothGrowth$dose == 0.5], paired = FALSE, var.equal = TRUE)$conf

t.test(ToothGrowth$len[ToothGrowth$dose == 2.0], ToothGrowth$len[ToothGrowth$dose == 1.0], paired = FALSE, var.equal = TRUE)$conf

t.test(ToothGrowth$len[ToothGrowth$dose == 2.0], ToothGrowth$len[ToothGrowth$dose == 0.5], paired = FALSE, var.equal = TRUE)$conf

```

Conclusion

From the result, the confidence intervals for three comparisons are all above zero, therefore, we can conclude that the larger the supp dose, the longer the tooth.

